opXRD: Open Experimental Powder X-ray Diffraction Database

### 2024-12-19 chapters/01_introduction.tex
Updated research projects in overview table

% VISUAL ABSTRACT FIGURE
\begin{figure}[!b]
    \centering
    \includegraphics[width=1.0\linewidth]{figures/figure1_v5.pdf}
    \caption{Overview}
    \label{fig:overview}
\end{figure}

@@

However, even small labeled and also unlabeled experimental diffractograms hold significant value in practice. First of all, labeled data can be used to test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset\cite{Armbruster2015}. \\

>>>>

% VISUAL ABSTRACT FIGURE
\begin{figure}[!b]
    \centering
    \includegraphics[width=1.0\linewidth]{figures/figure1_v5.pdf}
    \caption{Experimental powder x-ray diffraction (pXRD) patterns from several contributors are collected in the \textit{opXRD} database. The proposed open-access database of experimental data aims to support each step in the pXRD-related machine learning workflow by informing better physics simulations, supplying model training data, and providing a foundation for realistic performance evaluations.}
    \label{fig:overview}
\end{figure}

@@

However, even small labeled and also unlabeled datasets of experimental diffractograms hold significant value in practice. First of all, labeled data can be used to test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset\cite{Armbruster2015}. \\

### 2024-12-19 chapters/03_existing_datasets.tex
Updated research projects in overview table

To contextualize opXRD within the current environment of experimental powder diffraction data, the below list provides an overview of the largest crystal structure databases that offer access to experimental powder diffraction data. For an overview of these databases refer to Table$\eqref{tab:exp_databases}$ below. In this table, the column ``O.A.'' indicates whether or not the database is open access. The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.

\begin{table}[!htb]
\centering
\caption{Overview of experimental powder diffraction databases}
\label{tab:exp_databases}
\scalebox{0.84}{
\begin{adjustbox}{center}
\begin{NiceTabular}{@{}llllllll@{}}
\toprule
\textbf{Name} & \textbf{No. patterns} & \textbf{O.A.} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Atom coords.} & \textbf{Year est.} \\
\midrule
Linus Pauling file                     & 21,700                    & \xmark          & \cmark        & \cmark        & \cmark        & \cmark          & 2002            \\
Powder Diffraction File \tabularnote[a)]{The PDF lists the Material Platform for Data Science (MPDS) as a database source. Since the MPDS is hosted by the Pauling File project, there is likely significant overlap in the experimental patterns available in the PDF and the Linus Pauling File.} & 20,800 & \xmark & \cmark & \cmark & \cmark & \partialcheck{52} & 1941 \\
RRUFF                                  & 1290                      & \cmark          & \cmark        & \cmark        & \cmark        & \xmark          & 2006            \\
Crystallography Open Database          & 1052                      & \cmark          & \cmark        & \cmark        & \partialcheck{85}        & \partialcheck{85}          & 2003            \\
PowBase                                & 169                       & \cmark          & \cmark        & \xmark        & \xmark        & \xmark          & 1999            \\
\bottomrule
\end{NiceTabular}
\end{adjustbox}}
\end{table}

>>>>

To contextualize opXRD within the current environment of experimental powder diffraction data, the below list provides an overview of the largest crystal structure databases that offer access to experimental powder diffraction data. For an overview of these databases refer to Table~$\eqref{tab:exp_databases}$ below.

\begin{table}[!htb]
\centering
\caption{\footnotesize Overview of experimental powder diffraction databases: The column ``O.A.'' indicates whether or not the database is open access. The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.}
\label{tab:exp_databases}
\scalebox{0.84}{
\begin{adjustbox}{center}
\begin{NiceTabular}{@{}llllllll@{}}
\toprule
\textbf{Name} & \textbf{No. patterns} & \textbf{O.A.} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Atom coords.} & \textbf{Year est.} \\
\midrule
Linus Pauling file                     & 21,700                    & \xmark          & \cmark        & \cmark        & \cmark        & \cmark          & 2002            \\
Powder Diffraction File \tabularnote[a)]{The PDF lists the Material Platform for Data Science (MPDS) as a database source. Since the MPDS is hosted by the Pauling File project, there is likely significant overlap in the experimental patterns available in the PDF and the Linus Pauling File.} & 20,800 & \xmark & \cmark & \cmark & \cmark & \partialcheck{52} & 1941 \\
RRUFF                                  & 1290                      & \cmark          & \cmark        & \cmark        & \cmark        & \xmark          & 2006            \\
Crystallography Open Database          & 1052                      & \cmark          & \cmark        & \cmark        & \partialcheck{85}        & \partialcheck{85}          & 2003            \\
PowBase                                & 169                       & \cmark          & \cmark        & \xmark        & \xmark        & \xmark          & 1999            \\
\bottomrule
\end{NiceTabular}
\end{adjustbox}}
\end{table}

### 2024-12-19 chapters/04_our_dataset.tex
Updated research projects in overview table

In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiments, some of them labeled with corresponding structural information. The following paragraphs detail the contribution of each research group. Most data was collected using \ce{Cu} radiaton sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$. We filtered the submitted datasets to exclude patterns with invalid features such as the pattern containing only one unique recorded angle or negative angles, the pattern containing less than 50 recorded angles total, or all intensities being zero. \\

The below Table$\eqref{tab:merged}$ provides an overview of the opXRD database. The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.

\begin{table}[!htb]
\centering
\caption{The opXRD database}
\label{tab:merged}
\scalebox{0.81}{
\begin{adjustbox}{center}
\begin{tabular}{@{}lllllll@{}}
\toprule
\textbf{Institution} & \textbf{No. patterns} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Atom coords.} & \textbf{Research Project} \\
\midrule
CNRS        &  1052          & \cmark   & \partialcheck{85}   & \cmark   &  \partialcheck{85}   & Extracted from the COD \\ 
HKUST(GZ)   &  21           & \cmark   & \cmark   & \cmark   & \cmark   & Phase identification reference database \\ 
USC         &  338          & \cmark   & \cmark   & \partialcheck{90}   & \xmark   & Study of CuNi and CuAl alloys          \\ 
EMPA        &  790          & \cmark   & \partialcheck{63}   & \xmark   & \xmark   & Tin halide perovskites, Zn-V-N library \\ 
INT         &  19,780        & \xmark   & \xmark   & \xmark   & \xmark   & Compilation of various projects        \\ 
IKFT        &  64           & \xmark   & \xmark   & \xmark   & \xmark   & Commercial catalysts, bulk materials   \\ 
LBNL        &  n/a          & \xmark   & \xmark   & \xmark   & \xmark   & Time series recordings of Perovskites  \\ 
\bottomrule
\end{tabular}
\end{adjustbox}}
\end{table}

@@

A small-scale experimental powder X-ray database called the X-Ray Phase Identification Public Experimental Dataset (XRed) was established over the past two years.\footnote{XRed can be accessed on GitHub at \url{https://github.com/WPEM/XRED}.} The primary objective of this database is to support the development of intelligent phase identification technology, serving as a foundation for data collection for future large-scale machine learning calculations. \\

The data in XRed were generated in our laboratory using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers, all utilizing \ce{Cu} X-ray sources. Hundreds of pXRD patterns have been obtained through XRD refinement, enabling the determination of detailed lattice constants, atomic sizes, and other structural parameters. These patterns are labeled with CIF files that document the refined structures. The dataset is organized by elemental systems and includes original experimental files ranging from single-phase to five-phase mixtures, as well as mixtures studied for different tasks. \\

>>>>

In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiments, some of them labeled with corresponding structural information. The following paragraphs detail the contribution of each research group. Most data was collected using \ce{Cu} radiaton sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$. We filtered the submitted datasets to exclude patterns with invalid features such as the pattern containing only one unique recorded angle or negative angles, the pattern containing less than 50 recorded angles total, or all intensities being zero. The below Table$\eqref{tab:merged}$ provides an overview of the opXRD database.

\begin{table}[!htb]
\centering
\caption{\footnotesize The opXRD database: The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.}
\label{tab:merged}
\scalebox{0.79}{
\begin{adjustbox}{center}
\begin{tabular}{@{}lllllll@{}}
\toprule
\textbf{Institution} & \textbf{No. patterns} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Atom coords.} & \textbf{Research Project} \\
\midrule
CNRS        &  1052          & \cmark   & \partialcheck{85}   & \cmark   &  \partialcheck{85}   & Diffraction data extracted from COD \\ 
HKUST(GZ)   &  21+499          & \cmark   & \cmark   & \cmark   & \cmark   & Phase/structure identification dataset  \\ 
USC         &  338          & \cmark   & \cmark   & \partialcheck{90}   & \xmark   & Study of CuNi and CuAl alloys          \\ 
EMPA        &  790          & \cmark   & \partialcheck{63}   & \xmark   & \xmark   & Tin halide perovskites, Zn-V-N library \\ 
INT         &  19,780        & \xmark   & \xmark   & \xmark   & \xmark   & Compilation of various projects        \\ 
IKFT        &  64           & \xmark   & \xmark   & \xmark   & \xmark   & Commercial catalysts, bulk materials   \\ 
LBNL        &  n/a          & \xmark   & \xmark   & \xmark   & \xmark   & Perovskites precursors, Mn-Sb-O system  \\ 
\bottomrule
\end{tabular}
\end{adjustbox}}
\end{table}

@@

A small-scale experimental powder X-ray database called the X-Ray Phase Identification Public Experimental Dataset (XRed) was established over the past two years.\footnote{XRed can be accessed on GitHub at \url{https://github.com/WPEM/XRED}.} The primary objective of this database is to support the development of intelligent phase identification technology, serving as a foundation for data collection for future large-scale machine learning calculations. The data in XRed were generated using instruments such as the Empyrean 3.0, Aeris, and Bruker D8 Advance diffractometers, all utilizing \ce{Cu} X-ray sources. Hundreds of pXRD patterns have been obtained through XRD refinement, enabling the determination of detailed lattice constants, atomic sizes, and other structural parameters. These patterns are labeled with CIF files that document the refined structures. The dataset is organized by elemental systems and includes original experimental files ranging from single-phase to five-phase mixtures, as well as mixtures studied for different tasks. \\

Another experimental database integrated with opXRD is the "collected database", which comprises powder XRD data paired with corresponding crystal structures. The data were sourced from publicly available open-access publications and our collaborating institutions, with full authorization for research use. This database surpasses Xred in size, containing 499 entries. However, these entries have not undergone individual verification via pattern refinement. In most cases, the XRD patterns are labeled based on powder diffraction file cards rather than detailed pattern refinements.\\

### 2024-12-19 chapters/05_how_to_use.tex
Updated research projects in overview table

The opXRD database is hosted on Zenodo\footnote{The dataset can be found at \url{https://zenodo.org/records/14278656}.} and can be downloaded by any user without any barriers or restrictions. The database comes in two files, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''. The latter contains the in-situ data which contains highly correlated patterns recorded through time series measurements. Patterns belonging to time series measurements have filenames that indicate the measurement series they belong to as well as their order in that series. \\

>>>>

The opXRD database is hosted on Zenodo\footnote{The dataset can be found at \url{https://zenodo.org/records/14278656}.} and can be downloaded by any user without any barriers or restrictions. The database comes in two zip archives, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''. The latter contains the in-situ data which contains highly correlated patterns recorded through time series measurements. Within the .zip archives patterns are saved as .json files grouped in folders indicating the contributing institution. Each .json file contains a pattern recorded from an X-ray diffraction experiment. If available, the composition and structure of the investigated sample and experiment conditions are also included in this file. Patterns belonging to time series measurements are labeled with filenames that indicate the measurement series they belong to as well as their order in that series.

### 2024-12-19 main.tex
Recommended cuts to abstract
Updated research projects in overview table

\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\textcolor{green!80!black}{\ding{52}}}
\newcommand{\xmark}{\textcolor{red}{\ding{53}}}
\newcommand{\partialcheck}[1]{\textcolor{orange}{\ding{52}(#1\%)}}

\date{}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\DeclareSIUnit{\myeuro}{\text{\euro}}
\DeclareSIUnit\angstrom{\text {Å}}
\usepackage[final]{changes}

@@

\author[\ITI,\INT]{Daniel Hollarek}
\author[\ITI,\INT]{Henrik Schopmans}
\author[\INT]{Jona Östreicher}
\author[\ITI,\INT]{Jonas Teufel}
\author[\HKUSTGZ]{Bin Cao}
\author[\INT]{Simon Schweidler}
\author[\USC]{Adie Alwen}
\author[\LBLMF]{Mriganka Singh}
\author[\LBLMF,\LBLALS]{Tim Kodalle}
\author[\Hoffmann]{Hanlin Hu}
\author[\LBLCS]{Gregoire Heymans}
\author[\EMPA]{Alexander Wieczorek}
\author[\EMPA]{Siarhei Zhuk}
\author[\CPT]{Arthur Hardiagon}
\author[\KFUPMMSE,\KFUPMIRCIMR]{Maged Abdelsamie}
\author[\HKUSTGZ]{Longhan Zhang}
\author[\HKUSTGZ]{Tong-yi Zhang}
\author[\USC]{Andrea M. Hodge}
\author[\FZJIEMD]{Ruth Schwaiger}
\author[\INT]{Ben Breitung}
\author[\EMPA]{Sebastian Siol}
\author[\KITICRT]{Moritz Wolf}
\author[\CPT]{François-Xavier Coudert}
\author[\LBLMF]{Carolin M. Sutter-Fella}
\author[\ITI,\INT]{Pascal Friederich}

\renewcommand\Affilfont{\small}

\affil[\ITI]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}
\affil[\INT]{Institute of Nanotechnology, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}
\affil[\HKUSTGZ]{Guangzhou Municipal Key Laboratory of Materials Informatics, Hong Kong University of Science and Technology (Guangzhou), Guangzhou 511400, China}
\affil[\USC]{Department of Chemical Engineering and Materials Science, University of Southern California, Los Angeles CA 90089, USA}
\affil[\LBLMF]{Lawrence Berkeley National Laboratory, Molecular Foundry Division, Berkeley 94720 CA, USA}
\affil[\LBLALS]{Lawrence Berkeley National Laboratory, Advanced Light Source, Berkeley 94720 CA, USA}
\affil[\Hoffmann]{Hoffmann Institute of Advanced Materials, Shenzhen Polytechnic, Shenzhen 518055, China}
\affil[\LBLCS]{Lawrence Berkeley National Laboratory, Chemical Sciences Division, Berkeley 94720 CA, USA}
\affil[\EMPA]{Empa–Swiss Federal Laboratories for Materials Science and Technology, 8600 Dübendorf, Switzerland}
\affil[\KITICRT]{Engler-Bunte-Institut \& Institute of Catalysis Research and Technology, Karlsruhe Institute of Technology, Karlsruhe, Germany}
\affil[\CPT]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France}
\affil[\KFUPMMSE]{Material Science and Engineering Department, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia}
\affil[\KFUPMIRCIMR]{Interdisciplinary Research Center for Intelligent Manufacturing and Robotics, KFUPM, Dhahran 31261, Saudi Arabia}
\affil[\FZJIEMD]{Institute of Energy Materials and Devices, Forschungszentrum Juelich GmbH, 52425 Juelich, Germany}

@@

\maketitle
\begin{abstract}
\input{chapters/00_abstract}
\end{abstract}

>>>>

\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\textcolor{green!80!black}{\ding{52}}}
\newcommand{\xmark}{\textcolor{red}{\ding{53}}}
\newcommand{\partialcheck}[1]{\textcolor{orange}{\ding{52}(#1\%)}}
\captionsetup[table]{
  font=footnotesize
}

\date{}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\renewcommand\Authfont{\small\raggedright\itshape}  
\renewcommand\Affilfont{\normalfont\small}

\DeclareSIUnit{\myeuro}{\text{\euro}}
\DeclareSIUnit\angstrom{\text {Å}}
\usepackage[final]{changes}

@@

\author[\ITI,\INT]{Daniel Hollarek}
\author[\ITI,\INT]{Henrik Schopmans}
\author[\ITI,\INT]{Jona Östreicher}
\author[\ITI,\INT]{Jonas Teufel}
\author[\HKUSTGZ]{Bin Cao}
\author[\INT]{Simon Schweidler}
\author[\USC]{Adie Alwen}
\author[\LBLMF]{Mriganka Singh}
\author[\LBLMF,\LBLALS]{Tim Kodalle}
\author[\Hoffmann]{Hanlin Hu}
\author[\LBLCS]{Gregoire Heymans}
\author[\EMPA]{Alexander Wieczorek}
\author[\EMPA]{Siarhei Zhuk}
\author[\CPT]{Arthur Hardiagon}
\author[\KFUPMMSE,\KFUPMIRCIMR]{Maged Abdelsamie}
\author[\HKUSTGZ]{Longhan Zhang}
\author[\HKUSTGZ]{Tong-yi Zhang}
\author[\USC]{Andrea M. Hodge}
\author[\FZJIEMD]{Ruth Schwaiger}
\author[\INT]{Ben Breitung}
\author[\EMPA]{Sebastian Siol}
\author[\KITICRT]{Moritz Wolf}
\author[\CPT]{François-Xavier Coudert}
\author[\LBLMF]{Carolin M. Sutter-Fella}
\author[\ITI,\INT]{Pascal Friederich$^\ast$}

\affil[\ITI]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany.  E-mail: pascal.friederich@kit.edu}
\affil[\INT]{Institute of Nanotechnology, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}
\affil[\HKUSTGZ]{Guangzhou Municipal Key Laboratory of Materials Informatics, Hong Kong University of Science and Technology (Guangzhou), Guangzhou 511400, China}
\affil[\USC]{Department of Chemical Engineering and Materials Science, University of Southern California, Los Angeles CA 90089, USA}
\affil[\LBLMF]{Lawrence Berkeley National Laboratory, Molecular Foundry Division, Berkeley 94720 CA, USA}
\affil[\LBLALS]{Lawrence Berkeley National Laboratory, Advanced Light Source, Berkeley 94720 CA, USA}
\affil[\Hoffmann]{Hoffmann Institute of Advanced Materials, Shenzhen Polytechnic, Shenzhen 518055, China}
\affil[\LBLCS]{Lawrence Berkeley National Laboratory, Chemical Sciences Division, Berkeley 94720 CA, USA}
\affil[\EMPA]{Empa–Swiss Federal Laboratories for Materials Science and Technology, 8600 Dübendorf, Switzerland}
\affil[\KITICRT]{Engler-Bunte-Institut \& Institute of Catalysis Research and Technology, Karlsruhe Institute of Technology, Karlsruhe, Germany}
\affil[\CPT]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France}
\affil[\KFUPMMSE]{Material Science and Engineering Department, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia}
\affil[\KFUPMIRCIMR]{Interdisciplinary Research Center for Intelligent Manufacturing and Robotics, KFUPM, Dhahran 31261, Saudi Arabia}
\affil[\FZJIEMD]{Institute of Energy Materials and Devices, Forschungszentrum Juelich GmbH, 52425 Juelich, Germany}

@@

\maketitle
\textbf{Keywords}: powder x-ray diffraction, machine learning, open-access database, crystal structure determination, phase identification

\begin{abstract}
\input{chapters/00_abstract}
\end{abstract}
